Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-16 prop_iop.tex
prop_iop.tex \section{Why Empty Core?} Updating...

\begin{definition} Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$: 
\begin{itemize}
\item A \emph{core} of $\bar{\mu}$ is the set $\mathrm{core}\left(\bar{\mu}\right)=\set{\pmeas:\events\rightarrow[0,1]}{\forall P\in\events.~\pmeas\left(P\right)\in\bar{\mu}\left(P\right)}$. 
\item $\bar{\mu}$ is called convex if $\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$
for all commuting $P_{0},P_{1}\in\events$. 
\end{itemize}
\end{definition}

@@

\begin{example}[Three-dimensional quantum three-interval-valued
probability measure]\label{ex:three-dimensional-three-value} Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and 
\begin{equation}\eqalign{
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})=\imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})=\necess\textrm{ ,}\\
\bar{\mu}(P) = \unknown\qquad\textrm{otherwise.}
}\end{equation}
It is straightforward to verify that $\bar{\mu}$ is a convex quantum
interval-valued probability measure. Now assume there exists a real-valued
quantum probability measure $\mu$ such that $\mu(P)\in\bar{\mu}(P)$
for every event (projection)~$P$. We derive a contradiction as follows.
Suppose $\mu(P)\in\bar{\mu}(P)$ then we must have $\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0$.
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. However, no pure state~$\ket{\phi}$
can satisfy $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$; a contradiction.
\end{example}

\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figure2} 
\par\end{centering}
\caption{\label{fig:three-dimensional-4-value}This figure illustrates cases
1 and 2 of example~\ref{ex:three-dimensional-4-value} plotted in
$\mathbb{R}^{3}$. The red and green dotted vectors are $\ket{0}$
and $\ket{\ps}$ respectively. All possible real vectors of the subspaces
$\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\ps_{\theta,\gamma}^{\perp}}$
are drawn in the red and green circles, respectively. Within the circles,
a dotted vector~$\ket{\psi}$ means $\bar{\mu}(\proj{\psi})=\likely$;
otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is
a generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red
and green solid vectors are normalized $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively,
where $\times$ is the usual cross product in $\mathbb{R}^{3}$.}
\end{figure}

@@

Notice that in the previous section,

@@

\subsection{From Quantum to Classical}

@@

Looking closely at the example above, we might think that the probability
measure $\bar{\mu}$ is induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or we might think it is induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$,
or yet we might think it is induced by the mixed state $\mathbb{1}/3$
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$.
Each of these possibilities is compatible with some but not all of
the observations. All is not lost however: if the observations are
made more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous convex probability measure by adding more
precise intervals.

\begin{example}[Three-dimensional quantum four-interval-valued 
  probability measure]\label{ex:three-dimensional-4-value}
  Given a three dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.  Let \emph{
  }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$, and
  $\ket{\minus}=\left(\ket{0}-\ket{1}\right) / \sqrt{2}$. The definition of
  $\bar{\mu}$ below refers to figure~\ref{fig:three-dimensional-4-value}
  which plots the 1-dimensional projectors:
\begin{enumerate}
\item Let:
\begin{equation}\eqalign{
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess\textrm{ ,}
}\end{equation}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green dotted
vectors, respectively.
\item The red and green circles are are the states orthogonal to
  $\ket{0}$ and $\ket{\ps}$, respectively, and can be parametrized as
  $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
  and
  $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\minus}+\cos\theta\ket{2}$,
  where $0\le\theta\le\pi / 2$ and $0\le\gamma<2\pi$. The dotted
  half of those states need special treatment, i.e., whenever
  $0\le\theta<\pi / 2$ and $0\le\gamma<\pi$, we define
\begin{equation}\eqalign{
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\likely\textrm{ ,}\\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right)=
  \bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\unlikely\textrm{ .}
}\end{equation}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\begin{equation}
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\end{equation}
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\mathbb{1} / 3$, but not by $\ket{1}$. \end{example}

>>>>

\begin{definition} Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$: 
\begin{itemize}
\item A \emph{core} of $\bar{\mu}$ is the set $\mathrm{core}\left(\bar{\mu}\right)=\set{\pmeas:\events\rightarrow[0,1]}{\forall P\in\events.~\pmeas\left(P\right)\in\bar{\mu}\left(P\right)}$. 
\item $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
for all commuting $P_{0},P_{1}\in\events$. 
\end{itemize}
\end{definition}

@@

\begin{example}[Three-dimensional quantum three-interval-valued probability
measure]\label{ex:three-dimensional-three-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and 
\begin{equation}\eqalign{ 
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})=\imposs\textrm{ ,}\\ 
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})=\necess\textrm{ ,}\\ 
\bar{\mu}(P) = \unknown\qquad\textrm{otherwise.} 
}\end{equation}
It is straightforward to verify that $\bar{\mu}$ is a convex quantum
interval-valued probability measure. Now assume there exists a real-valued
quantum probability measure $\mu$ such that $\mu(P)\in\bar{\mu}(P)$
for every event (projection)~$P$. We derive a contradiction as follows.
Suppose $\mu(P)\in\bar{\mu}(P)$ then we must have $\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0$.
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. However, no pure state~$\ket{\phi}$
can satisfy $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$; a contradiction.
\end{example}

@@

The possibility that a quantum interval-valued probability measure
has an empty core departs from the classical cases. In order to understand
its meaning, we will first study how to recover the non-empty core
by adding conditions. Then, we will consider having an empty core
as an legitimate result, and understand its consequence.

@@

One possibility is that the definition of quantum convexity is too
weak to guarantee an convex quantum interval-valued probability measure
has an non-empty core. However, the product of non-commuting projectors
is not a projector so that equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
cannot be meaningful when $P_{0}$ and $P_{1}$ are non-commuting.

Classically, equation~(\ref{eq:QuantumInterval-valuedProbability-Convex})
is a special case of 
\begin{equation}
\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)+\sum_{I\subseteq\left\{ 1,\ldots,N\right\} \textrm{ and }\left|I\right|\textrm{ is odd}}\bar{\mu}\left(\bigcap_{i\in I}E_{i}\right)\subseteq\sum_{I\subseteq\left\{ 1,\ldots,N\right\} \textrm{ and }\left|I\right|\textrm{ is even}}\bar{\mu}\left(\bigcap_{i\in I}E_{i}\right)\textrm{ .}\label{eq:Dempster-Shafer-game}
\end{equation}
If a classical interval-valued probability measure~$\bar{\mu}$ satisfies
equation~(\ref{eq:Dempster-Shafer-game}) for all set of event~$\left\{ E_{i}\right\} _{i=1}^{N}\subseteq2^{\Omega}$,
then the left-end point of $\bar{\mu}$ is called a Dempster-Shafer
belief function. Equation~(\ref{eq:Dempster-Shafer-game}) could
be extended to the quantum case, and consider a quantum interval-valued
probability measure satisfying 
\begin{equation}
\bar{\mu}\left(\bigvee_{i=1}^{N}P_{i}\right)+\sum_{I\subseteq\left\{ 1,\ldots,N\right\} \textrm{ and }\left|I\right|\textrm{ is odd}}\bar{\mu}\left(\prod_{i\in I}P_{i}\right)\subseteq\sum_{I\subseteq\left\{ 1,\ldots,N\right\} \textrm{ and }\left|I\right|\textrm{ is even}}\bar{\mu}\left(\prod_{i\in I}P_{i}\right)\textrm{ .}\label{eq:Dempster-Shafer-game-quantum}
\end{equation}
for all set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events$.
However, example~\ref{ex:three-dimensional-three-value} satisfies
equation~(\ref{eq:Dempster-Shafer-game-quantum}) as well. In another
word, this kind of conditions cannot eliminate the situation of empty
core.

\subsection{Restricting the Event Space}

@@

Looking closely at the example above, we might think that the probability
measure $\bar{\mu}$ is induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or we might think it is induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$,
or yet we might think it is induced by the mixed state $\mathbb{1}/3$
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$.
Each of these possibilities is compatible with some but not all of
the observations. All is not lost however: if the observations are
made more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous convex probability measure by adding more
precise intervals.
\begin{figure}
\begin{centering}
\includegraphics[scale=0.38]{figure2} 
\par\end{centering}
\caption{\label{fig:three-dimensional-4-value}This figure illustrates cases
1 and 2 of example~\ref{ex:three-dimensional-4-value} plotted in
$\mathbb{R}^{3}$. The red and green dotted vectors are $\ket{0}$
and $\ket{\ps}$ respectively. All possible real vectors of the subspaces
$\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\ps_{\theta,\gamma}^{\perp}}$
are drawn in the red and green circles, respectively. Within the circles,
a dotted vector~$\ket{\psi}$ means $\bar{\mu}(\proj{\psi})=\likely$;
otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is
a generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red
and green solid vectors are normalized $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively,
where $\times$ is the usual cross product in $\mathbb{R}^{3}$.}
\end{figure}

\begin{example}[Three-dimensional quantum four-interval-valued probability
measure]\label{ex:three-dimensional-4-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, and $\ket{\minus}=\left(\ket{0}-\ket{1}\right)/\sqrt{2}$.
The definition of $\bar{\mu}$ below refers to figure~\ref{fig:three-dimensional-4-value}
which plots the 1-dimensional projectors: 
\begin{enumerate}
\item Let:
\begin{equation}\eqalign{
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess\textrm{ ,}
}\end{equation}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green dotted
vectors, respectively.
\item The red and green circles are are the states orthogonal to
  $\ket{0}$ and $\ket{\ps}$, respectively, and can be parametrized as
  $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
  and
  $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\minus}+\cos\theta\ket{2}$,
  where $0\le\theta\le\pi / 2$ and $0\le\gamma<2\pi$. The dotted
  half of those states need special treatment, i.e., whenever
  $0\le\theta<\pi / 2$ and $0\le\gamma<\pi$, we define
\begin{equation}\eqalign{
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\likely\textrm{ ,}\\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right)=
  \bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\unlikely\textrm{ .}
}\end{equation}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\begin{equation}
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\end{equation}
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\mathbb{1} / 3$, but not by $\ket{1}$. \end{example}
